Does the COMPAS Needle Always Point Towards Equity? Finding Fairness in the COMPAS Risk Assessment Algorithm: A Case Study

### 2022-05-03 Paper_Draft2/Draft2.Rmd
fixing image bug

The data we are using for this project is the COMPAS General Recidivism Risk Scores data set from the AI Fairness 360 (AIF360) toolkit [@aif360-oct-2018], which does the same initial pre-processing as ProPublica. The raw data has 6,167 rows and each row represents an arrest charge for a defendant. AIF360’s COMPAS data includes the defendant's age, race, sex, prior charges, what they were charged with, and whether or not the defendant ultimately recidivated within a two-year period after their arrest. For the purposes of our project, which endeavors to evaluate the differing effects of the COMPAS algorithm on white defendants versus Black defendants, we have filtered the data to only include individuals whose race is listed as Caucasian or African-American. Our data therefore has 5,273 rows (Figure \ref{fig:table snip}), with the distributions of age (Figure \ref{fig:age plot}), prior charges (Figure \ref{fig:prior plot}), and two-year recidivism by race (Figure \ref{fig:recidivism mosaic}) and (Table \ref{tab:recid table}) shown below.

@@

```{r recidivism mosaic, echo=FALSE, fig.align = 'center', fig.cap = 'When we divide the data into Black and white defendants, we can see that Black defendants recidivate more than white defendants and Black defendants are more likely to recidivate than not recidivate. 39.14% of white defendants did recidivate within two years compared to 52.35% of black defendants. We can also see that there are more Black defendants in the dataset overall.', out.width= "100%"}
knitr::include_graphics("../images/recidivism_mosaic.png")
```

>>>>

The data we are using for this project is the COMPAS General Recidivism Risk Scores data set from the AI Fairness 360 (AIF360) toolkit [@aif360-oct-2018], which does the same initial pre-processing as ProPublica. The raw data has 6,167 rows and each row represents an arrest charge for a defendant. AIF360’s COMPAS data includes the defendant's age, race, sex, prior charges, what they were charged with, and whether or not the defendant ultimately recidivated within a two-year period after their arrest. For the purposes of our project, which endeavors to evaluate the differing effects of the COMPAS algorithm on white defendants versus Black defendants, we have filtered the data to only include individuals whose race is listed as Caucasian or African-American. Our data therefore has 5,273 rows (Figure \ref{fig:table snip}), with the distributions of age (Figure \ref{fig:age plot}), prior charges (Figure \ref{fig:prior plot}), and two-year recidivism by race (Figure \ref{fig:recidivism mosaic} and Table \ref{tab:recid table}) shown below.

@@

```{r recidivism mosaic, echo=FALSE, fig.align = 'center', fig.cap = 'When we divide the data into Black and white defendants, we can see that Black defendants recidivate more than white defendants and Black defendants are more likely to recidivate than not recidivate. 39.14\\% of white defendants did recidivate within two years compared to 52.35\\% of Black defendants. We can also see that there are more Black defendants in the dataset overall.', out.width= "100%"}
knitr::include_graphics("../images/recidivism_mosaic.png")
```
